---
title: "Intro_GeoSpatial_Analysis"
author: "Mauro Mardones"
date: "2023-03-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(sf)
library(ggplot2)
library(leaflet)
library(scales)
library(ggmap)
library(dplyr)
```


```{r}
## Read in shapefile using sf
ak_regions <- read_sf("~/training_mardones/data/ak_regions_simp.shp")

plot(ak_regions)  
```
```{r}
class(ak_regions)
```

Coordinate Reference System

Every sf object needs a coordinate reference system (or crs) defined in order to work with it correctly. A coordinate reference system contains both a datum and a projection. The datum is how you georeference your points (in 3 dimensions!) onto a spheroid. The projection is how these points are mathematically transformed to represent the georeferenced point on a flat piece of paper. All coordinate reference systems require a datum. However, some coordinate reference systems are “unprojected” (also called geographic coordinate systems). Coordinates in latitude/longitude use a geographic (unprojected) coordinate system. One of the most commonly used geographic coordinate systems is WGS 1984.

ESRI has a blog post that explains these concepts in more detail with very helpful diagrams and examples.

You can view what crs is set by using the function `st_crs`
```{r}
st_crs(ak_regions)
```


This is pretty confusing looking. Without getting into the details, that long string says that this data has a greographic coordinate system (WGS84) with no projection. A convenient way to reference crs quickly is by using the EPSG code, a number that represents a standard projection and datum. You can check out a list of (lots!) of EPSG codes [here](https://spatialreference.org/ref/epsg/?page=1).

We will use several EPSG codes in this lesson. Here they are, along with their more readable names:

3338: Alaska Albers
4326: WGS84 (World Geodetic System 1984), used in GPS
3857: Pseudo-Mercator, used in Google Maps, OpenStreetMap, Bing, ArcGIS, ESRI
You will often need to transform your geospatial data from one coordinate system to another. The st_transform function does this quickly for us. You may have noticed the maps above looked wonky because of the dateline. We might want to set a different projection for this data so it plots nicer. A good one for Alaska is called the Alaska Albers projection, with an EPSG code of 3338.


```{r}
ak_regions_3338 <- ak_regions %>%
  st_transform(crs = 3338)

st_crs(ak_regions_3338)

```

```{r}
plot(ak_regions_3338)
```
Since sf objects are `data.frames`, they play nicely with packages in the tidyverse. Here are a couple of simple examples:

`select()`
```{r}
ak_regions_3338 %>%
  select(region)


ak_regions_3338 %>%
  filter(region == "Southeast")
```

## Join


You can also use the sf package to create spatial joins, useful for when you want to utilize two datasets together. As an example, let’s ask a question: how many people live in each of these Alaska regions?

We have some population data, but it gives the number of people by city, not by region. To determine the number of people per region we will need to:

read in the city data from a csv and turn it into an sf object
use a spatial join (st_join) to assign each city to a region
use group_by and summarize to calculate the total population by region
First, read in the population data as a regular data.frame. This data is derived from: Jeanette Clark, Sharis Ochs, Derek Strong, and National Historic Geographic Information System. 2018. Languages used in Alaskan households, 1990-2015. Knowledge Network for Biocomplexity. doi:10.5063/F11G0JHX. Unnecessary columns were removed and the most recent year of data was selected.



```{r}
pop <- read.csv("~/training_mardones/data/alaska_population.csv")
```



We can do this easily using the `st_as_sf` function, which takes as arguments the coordinates and the crs. The `remove = F` specification here ensures that when we create our `geometry` column, we retain our original lat lng columns, which we will need later for plotting. Although it isn’t said anywhere explicitly in the file, let’s assume that the coordinate system used to reference the latitude longitude coordinates is WGS84, which has a `crs number of 4236.


```{r}
pop_4326 <- st_as_sf(pop, 
                  coords = c('lng', 'lat'),
                  crs = 4326,
                  remove = F)

head(pop_4326)
```

```{r}
pop_3338 <- st_transform(pop_4326, crs = 3338)
```

```{r}
pop_joined <- st_join(pop_3338, ak_regions_3338, join = st_within)

head(pop_joined)
```

```{r}
pop_region <- pop_joined %>% 
  as.data.frame() %>% 
  group_by(region) %>% 
  summarise(total_pop = sum(population))

head(pop_region)
```
```{r}
pop_region_3338 <- left_join(ak_regions_3338, pop_region)

#plot to check
plot(pop_region_3338["total_pop"])
```



```{r}
pop_mgmt_338 <- pop_region_3338 %>% 
  group_by(mgmt_area) %>% 
  summarize(total_pop = sum(total_pop))

plot(pop_mgmt_338["total_pop"])
```
```{r}
pop_mgmt_3338 <- pop_region_3338 %>% 
  group_by(mgmt_area) %>% 
  summarize(total_pop = sum(total_pop), do_union = F)

plot(pop_mgmt_3338["total_pop"])
```

Writing the file

Save the spatial object to disk using `write_sf()` and specifying the filename. Writing your file with the extension .shp will assume an ESRI driver driver, but there are many other format options available.
```{r}
write_sf(pop_region_3338, "~/training_mardones/data/ak_regions_population.shp", delete_layer = TRUE)
```


## Vizualitation

```{r}
ggplot(pop_region_3338) +
  geom_sf(aes(fill = total_pop)) +
  theme_bw() +
  labs(fill = "Total Population") +
  scale_fill_continuous(low = "khaki", high =  "firebrick", labels = comma)
```

```{r}
rivers_3338 <- read_sf("~/training_mardones/data/ak_rivers_simp.shp")
st_crs(rivers_3338)
```

```{r}
ggplot() +
  geom_sf(data = pop_region_3338, aes(fill = total_pop)) +
  geom_sf(data = rivers_3338, aes(size = StrOrder), color = "black") +
  geom_sf(data = pop_3338, aes(), size = .5) +
  scale_size(range = c(0.01, 0.2), guide = F) +
  theme_bw() +
  labs(fill = "Total Population") +
  scale_fill_continuous(low = "khaki", high =  "firebrick", labels = comma)
```

